### 2025-06-10 intro.qmd
20250609 - learning goals and outcomes

## Learning Objectives {#sec-objectives}

This book aims to help readers accomplish the following learning objectives:

- Apply empirical inference and appreciate the value it provides over speculative supposition.
- Ask educated questions when confronted with decisions in the face of uncertainty.
- Understand human decision making, including common heuristics and cognitive biases and how to mitigate them analytically.
- Engage in critical thinking about causality, including devising plausible alternative explanations for observed effects.
- Understand causal inference including confounding, causal pathways, and counterfactuals.
- Think empirically about human behavior and performance.
- Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
- Apply basic skills in statistical programming using `R` to manipulate and summarize datasets and to conduct data analysis.
- Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.
- Use various analytical techniques for predicting the outcome of uncertain events, and for uncovering latent causes of patterns in observed data.
- Interpret findings from various statistical approaches and evaluate the accuracy of predictions.
- Engage in iterative problem-solving processes, refining analytical approaches based on feedback and outcomes, and adapting strategies accordingly.
- Communicate statistical findings and analyses in both written and oral formats, demonstrating proficiency in presenting complex information to diverse audiences.
- Make sense of big data.
- Use practical analytical skills that can be applied in future research and job settings.

>>>>

## Goals {#learningGoals}

The book has three overarching goals; to help students:

1. Appreciate the value of statistics for answering questions you find interesting
1. Think critically and be a critical consumer of information in society
1. Get excited about learning stats!

## Learning Outcomes {#learningOutcomes}

Through reading the book, readers will be able to:

1. Think critically and empirically about human behavior and performance
1. Make sense of big data
1. Use, evaluate, and interpret statistical analyses
1. Communicate data-driven insights

### How Readers Will Achieve the Learning Outcomes {#learningOutcomesHowAchieve}

Readers will achieve the learning outcomes through the following specific outcomes:

#### 1. Think critically and empirically about human behavior and performance {#learningOutcomesThinkCritically}

- Apply empirical inference and articulate its advantages over speculative supposition.
- Formulate research questions, hypotheses, and predictions.
- Provide well-reasoned guidance for decision-making under conditions of uncertainty.
- Describe common flaws in human judgment and decision making, including heuristics and cognitive biases, and explain how they can be mitigated analytically.
- Critically evaluate causal claims by generating plausible alternative explanations.
- Apply core concepts in causal inference, including confounding, causal pathways, and counterfactuals.
- Compare the strengths and limitations of human and machine approaches to prediction.

#### 2. Make sense of big data {#learningOutcomesBigData}

- Apply foundational skills in statistical programming using `R` to combine, manipulate, clean, and summarize large datasets.

#### 3. Use, evaluate, and interpret statistical analyses {#learningOutcomesAnalyses}

- Conduct data analyses using `R`, including appropriate application of statistical models.
- Critically evaluate the strengths and limitations of statistical models and methods used for predicting uncertain events.
- Apply analytical techniques to predict outcomes and uncover latent causes in observed data.
- Interpret results from statistical analyses and evaluate the accuracy of predictions.
- Engage in iterative problem-solving, refining analytical strategies based on results and feedback.
- Use practical analytical skills that can be applied in future research and job settings.

#### 4. Communicate data-driven insights {#learningOutcomesCommunication}

- Communicate statistical findings in writing.
- Identify characteristics of effective data visualizations.
- Create data visualizations.
